---
title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
date: "February 18, 2018"
geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty
      
tables: True
link-citations: True
urlcolor : blue

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---

<!--**DAN: right now we sometimes use the terminology "normal copula" and sometimes 
"Gaussin copula". Let's choose one and be consistent across the whole ms, including 
main text and supp mat. I am leaning toward "Gaussin copula" but could be convinced 
otherwise, it is just the consistency that I want to be sure we get to.-->

<!--**SG: I think normal copula would be better as "N" is the abbreviation for that but for "G" means gumbel not gaussian.
I made all "Gaussian" as "Normal" copula in both main and supmat.-->

updated on `r Sys.Date()`

```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
#seed<-101
#source("mtime.R") #A function needed for caching
```

# Front matter

*Affiliations:*

Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA

Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA

Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA

Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA. 

*Possible alternative titles:*

- A complete statistical description of dependence between variables for ecology and related fields

*Keywords:*

*Type of Article*

- Ideas and Perspectives

*The number of words in the abstract*

*The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)* 

*The number of words in each text box*

*The number of references*

*The number of figures, tables, and text boxes*

*The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent* 

\newpage
# Abstract

\newpage
# Introduction \label{Intro}

Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in the
finance and risk management. For many years, measuring and modeling dependencies has focused on the
Pearson correlation because of its ease in use. But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two random variables. Therefore, it is better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation. However, using copula-based approach as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number. The flexibility of copulas allows us to model more complex dependence structures by isolating its marginal from its joint distribution structures, with the capability in modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry. Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution. For example, in the upper panel of Fig. \ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, were shown which became dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we moved along the row. Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$). To implement this, we need to select candidate bivariate copulas from many existing parametric copula families. A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from. Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}. 

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously. If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them. Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \ref{SM-NPA} with three different statistics to verify our model selection based results. 

Basically, we want to address the following questions. \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? \textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure? And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Material&Methods}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections and are focused on possible causes and consequences of non-Normal copula structure in ecology and related field. Finally, we concluded in the \nameref{Discussion} section. 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption{Pedagogical figure showing BB1 copula with varying tail-dep difference but of same Kendall's tau along each row and with same tail-dep difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Material and methods \label{Material&Methods}

## Data \label{Data}  

We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each datasets used were given in \ref{SM-Data_detailed}.

\begin{figure}[!h]
\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs. ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs. contrasts for AVG.humerus phylogenetic traits. \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript). \label{fig_biv_raw_cop}}
\end{center}
\end{figure}


\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5). Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel. These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript). \label{fig_aphid_raw_cop}}
\end{center}
\end{figure}

## Statistical methods for detecting copula structure \label{Methods}

Question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered via statistical analysis of the
data described above, using both model selection and nonparametric approaches.

### Model selection approach \label{Model_selection}

A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) 
and Bayesian Information Criterion (BIC) was used to
help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}. 

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$ such as the soil C and N data, model selection involved several steps, 
most implemented using tools from the `VineCopula` package. Code combining these tools is in the
[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 

First, we produced ranks 
$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
smallest element, up to $n$, for the largest. 

Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
selection algorithms are ineffective if data could not be distinguished from independent data 
since most or all bivariate copula families considered include the independent copula in the 
family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic. If independence could be rejected ($0.05$ significance level), model selection proceeded.

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

Model selection methods will give the *relative* support of several models, but will not,
on their own, indicate whether any of the models selected is an objectively good fit (all 
models considered might poorly describe the data). To test this, and as the fourth step of
our procedure, we tested the goodness of fit of our best-AIC copula family using a 
bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
two tests, one based on a Cramer-von Mises statistic, and one based on a 
Kolmogorov-Smirnov statistic. For computational efficiency, 
an initial run using $100$ bootstraps was performed, and if
the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
families with that of the Normal family. The same was also done using BIC.
The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a Normal copula, both of which 
are 0. The third part of \hyperref[Intro]{\textbf{(Q1)}}
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a Normal copula, again 0.

For multivariate datasets, e.g., the aphid abundance or first flight data, 
each of which consists of time series from 10 locations, we carried out the same bivariate analysis
for all possible pairwise combinations of those locations and compared AIC-weights and tail-dependence of best-fit(minimum AIC) copula with that of the Normal copula. We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa. For example, for green-spruce aphid abundance data we found major lower tail dependence whereas leaf-curling plum aphid's first flight data had significant upper tail dependence at all location pairs (for details see Table \ref{tab_multivar_summary}).
<!--***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
pairwise-->
<!--**SG: done-->

<!--### Nonparametric analysis of tail dependence \label{NPA}-->

In addition to the model selection approach described above, **nonparametric approaches** (for details see Appendix \ref{SM-NPA}) to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust. 
 
<!--***DAN: There is either too little or too much here in this 
paragraph. The reader cannot get the main
idea, really, of what was done for these tests, from this paragraph alone - they will have to
look at the appendix. They need to be able to get the main idea. So you could either add detail,
or else take away detail and just say something like "We tested our statistics by testing whether 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed asymmetry of tail dependence 
known to be present in data simulated from various copulas (Appendix X)" and then stop there.
Then you would remove panels B-D from Fig. 1. At the moment I prefer this second option, but 
my preference is not strong, so if you prefer to keep the material and add detail that would be 
fine.-->
<!--**SG: Agreed with 2nd option and did things as necessary-->

<!--***DAN: I think the below material belongs in supp mat, because the reader can get the
main idea via the above text, and only some readers will be interested in the detailed 
algorithm. Please move it there, and then make the above links XYZ and ABC into live links
to the appropriate locations in the sup mat.

For bivariate datasets, surrogates were produced as follows.   

- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
study using the `iTau` function of the `VineCopula` package). 
- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
$\mathcal{B}(\theta_\tau)$. 
- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
indices $1,\ldots,n$). 
- Likewise permute $(y_1,\ldots,y_n)$
such that the permutation
$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
another permutation of $1,\ldots,n$).
- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$. 

Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
[`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 

Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.  
-->
<!--**SG: Done-->

<!--Theoretical models
Question 2 from the Introduction was explored using simple population models.-->


# Results \label{Results}

<!--## question 1

- Do datasets in ecology and related fields have non-Normal copula structure?
- Do they show tail dependence distinct from that of a Normal copula?
- In particular do they show asymmetric tail dependence?--> 

To answer \hyperref[Intro]{\textbf{(Q1)}}, we choose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.

### Example for bivariate case : soil C and N data \label{Ex_SoilCN}

```{r read_BivMS_soilCN_data,echo=F}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

As described in the Appendix \ref{SM-Data_detailed}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (E), respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

#### Model selection approach \label{Ex_SoilCN_Model_selection}

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`). We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit. The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data. 

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}. To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.

Although these results do convincingly show non-Normal copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}

As described in Appendix \ref{SM-NPA}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 Normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.

Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{fig_biv_raw_cop}(E). 
Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.

We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets that were plotted in Fig. \ref{fig_biv_raw_cop}. 

The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data. 
We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
arranged in ascending order. So, whenever those ranks were $>975$ or $<25$, that indicated a significantly ($95\%$ CI) different 
tail-dependence measures based on Normal-null hypothesis. 

For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. 
Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal. It was indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
(lower-upper) tail dependence and very high ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}}
(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}). 

```{r tab_bivar_summary, echo=F, results='markup'}

tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

```{r read_RES_aphid_count, echo=F}
RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
```

#### Model selection approach \label{Ex_count_Model_selection}

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis. 

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Fig. \ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \ref{SM-NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. 

Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig. \ref{SM-fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig. \ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach. Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 


```{r tab_multivar_summary, echo=F, results='markup'}

tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
knitr::kable(tab_multivar_summary, 
             format='pandoc',
             caption = "Summary of results for multivariate data-sets.\\label{tab_multivar_summary}",
             booktabs=TRUE)
```

We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations). First $11$ rows in Table \ref{tab_multivar_summary} represent
model selection results :
(1) total number of location pairs (excluding self-comparison) considered, 
(2) how many location pair copula are non-independent among them, 
(3) number of non-Normal copula as best fit (based on highest AIC-weight),
(4) average AIC-weights of best fit copula from all location pair,
(5) average AIC-weights of Normal copula from all location pair,
(6) number of location pair for which best fit copula have p-values > 0.01 
(both Cramer-von Mises and Kolmogorov-Smirnov method)
(7) lower and (8) upper quantile ($95\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair,
(9) lower and (10) upper quantile ($95\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair,
(11) number of location pair which showed stronger lower tail dependence than upper.

Last $6$ rows of Table \ref{tab_multivar_summary} summarized the non-parametric analysis results 
for $\Ps$ statistics : lower and upper quantile ($95\%$CI) of $\Ps_l$, $\Ps_u$, $\Ps_l-\Ps_u$ based on resampling
location pair.

Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
plankton data from UK seas also showed major lower tail dependent spatial synchrony 
($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.


<!--Question 2 analyses-->
## Some possible causes of non-Normal copula structure \label{Causes}

This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change? 
<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.--> 
<!--### The Moran effect and copula structure in populations-->
Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables. This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations. We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment and described our findings in Appendix \ref{SM-Cause_Moran}.
<!--
### Asymmetric sensitivity and copula structure in populations
Another model, distinct from the above, Dan and Shya to discuss-->
On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold. So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}). 

### Need to think about mechanisms for biogeochemical, evolutionary, and community data

<!--Question 3 analyses-->
## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}

This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data : we tested with four multivariate datasets and found results consistent with model selection and nonparametric approach (For details see Appendix \ref{SM-Consequence_skewness}). We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}). 

# Discussion \label{Discussion}



\newpage
# References


