---
author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
date: "February 18, 2018"
geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.8cm"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty
      
tables: True
link-citations: True
urlcolor : blue
indent : True

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---

<!--**DAN: right now we sometimes use the terminology "normal copula" and sometimes 
"Gaussin copula". Let's choose one and be consistent across the whole ms, including 
main text and supp mat. I am leaning toward "Gaussin copula" but could be convinced 
otherwise, it is just the consistency that I want to be sure we get to.-->

<!--**SG: I think normal copula would be better as "N" is the abbreviation for that but for "G" means gumbel not gaussian.
I made all "Gaussian" as "Normal" copula in both main and supmat.-->

updated on `r Sys.Date()`

```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
#seed<-101
#source("mtime.R") #A function needed for caching
```

# Front matter

*Affiliations:*

Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA

Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA

Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA

Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA

Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA. 

*Possible alternative titles:*

- A complete statistical description of dependence between variables for ecology and related fields

*Keywords:*

*Type of Article*

- Ideas and Perspectives

*The number of words in the abstract*

*The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)* 

*The number of words in each text box*

*The number of references*

*The number of figures, tables, and text boxes*

*The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent* 

\newpage
# Abstract

We introduce a new rank based "copula" approach to model dependencies between ecological variables. Our method could overcome the shortcomings of the linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, biogeochemistry and related field. We focus on three objectives (1) Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric
tail dependence? (2) If so, what are some possible causes of non-Normal copula structure? And (3) what are the
consequences of non-Normal copula structure and tail dependence for our understanding of biology and what are the applications? We present several examples with real data which show asymmetric tail dependence in bivariate as well as multivariate non-Normal copulae. Asymmetric tail dependence can illuminate some important phenomena like "spatial synchrony" and causal relationships between ecological variables (like species abundance and rainfall). We show that the "Moran-effect" and asymmetric sensitivity towards global environmental fluctuations could be the possible causes for non-Normal copula structures. We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and spatial Taylor's law slope values.  Our method is new and could have potential when exploring some other related phenomena like effects of climate change on extinction risks of metapopulations.

\newpage
# Introduction \label{Intro}

Very generally, all branches of ecology and related areas study relationships among environmental and biological 
variables. However, commonly used correlation- and regression-based 
approaches to studying 
such relationships are limited - it is well understood that common statistical
approaches provide only a very partial description of a relationship between variables. 
For instance, datasets showing markedly different 
relationships may nevertheless have the same correlation coefficient value:  
Fig. \ref{fig_cop_pedag1}A-B shows two such datasets. The variables of panel A 
(respectively, B) are principally 
related in the left (respectively, right) portions of their distributions, an asymmetric pattern of dependence in the distribution tails that may
have ecological significance, as discussed below, but is not captured by 
correlation. Panels C and D 
were obtained from A and B through monotonic transformation (Appendix X for details),
<!--***DAN: Dan to go back and write the appendix-->
which alters the Pearson, but not the Spearman correlation. Although correlation
is only one of the numerous metrics for 
studying relationships between variables, other commonly used metrics also 
provide partial descriptions only.

Well-developed statistical tools do exist, and are 
applied widely in other fields, that provide a demonstrably complete 
description of the dependence 
between two or more variables [@joe2014_dependence; @nelsen2006_copula]. 
The goal of this paper is to illustrate the potential for 
broader application of these
tools in ecology and related fields. The tools are based on 
the statistical concept of the *copula*, which we here 
introduce. From a data perspective, given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}, panel A or B), the associated copula
structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set of $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. See, 
for instance,
Fig. \ref{fig_cop_pedag1}E-F, which show the copula structure of 
A-B, respectively. Mathematically, if $(X,Y)$ is a continuous bivariate random
variable and the cummulative distribution functions (cdfs) of $X$ and $Y$ are 
$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the copula structure
of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
This random variable takes values in the unit square, 
$[0,1] \times [0,1]$, and has uniform marginal distributions. Densities
corresponding to the copulas associated 
with Fig. \ref{fig_cop_pedag1}E-F are in Fig.
\ref{fig_cop_pedag1}G-H, respectively. Strictly speaking, 
the copula is actually defined as the bivariate 
cdf of $(F_X(X),F_Y(Y))$ rather than as this random variable itself, 
but this makes little conceptual difference. See @nelsen2006_copula and @joe2014_dependence
for mathematical details. An important theorem of Abe Sklar [@sklar1959_theorem] 
provides the result that copulas give a unique and complete 
representation of dependence
between variable: the original joint distribution of $(X,Y)$ 
can be reconstructed from the 
copula and the marginal distributions. Copula theory is also well developed for
multivariate distributions.

One way we may expect a copula approach  
to better reflect ecological relationships has to do with Liebig's law of the
minimum. If, for instance, the growth of a plant depends on soil 
nitrogen, N, as well
as on other factors, data on soil nitrogen and resulting growth rates may be 
expected to look more like Fig. \ref{fig_cop_pedag1}A or E than like 
Fig. \ref{fig_cop_pedag1}B or F. Nitrogen 
will control plant growth, producing a strong
relationship, only when it is limiting, in the left portion of the N and 
plant-growth-rate distributions. 
Nitrogen will be less related or unrelated to growth rates in the right tails of 
the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E.

A second reason we may expect a copula-based approach to benefit 
ecological research is that such an approach may help illuminate 
causal mechanisms by which variables are related. For instance, if two species
X and Y are strong competitors, we may expect abundances of the two species
across quadrats or through time to be negatively related, as in 
Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
and Y is plotted on the vertical, then the plot in Fig. \ref{fig_cop_pedag1}I 
suggests X is the dominant 
competitor: when X is abundant, Y is necessarily rare because it is suppressed 
by X; whereas when X is rare, Y
may be abundant or may also be relatively rare due to 
factors other than X. Alternatively, Fig. \ref{fig_cop_pedag1}J suggests that Y
is the dominant competitor. 

A third reason we may expect a copula approach to benefit ecological
research is that prior work has demonstrated important copula 
stucture in the spatial dependence of environental 
variables [@Serinaldi2008; @She2018; @Goswami2018;
@li2013]. Such dependence tends to beget dependence across space
in fluctuations of populations or other ecosystem variables 
(this is called \emph{synchrony}) which are influenced by the environmental 
variables. This is called the \emph{Moran effect}. 
If copula structure is transmitted, 
possibly in modified form, 
from environmental to ecological variables, then we may expect important 
copula structure to be a common feature of dependencies through time 
between ecological variables measured in different locations. 
One may expect copula structure in environmental variables, in the form of 
accentuated tail-dependence, if
intense meteorological events are also spatially widespread.
We will explore this as well as other potential mechanisms for
copula structure in ecological variables. 

A final \emph{a priori} reason we may expect copula structure to be important
in ecology has to do with spatially aggregated or averaged quantities. 
Many ecological variables of applied importance 
depend on the spatial average or sum of local 
quantities. For instance, regional methane of CO$_2$ 
fluxes are the sum of local fluxes; 
and the total economic value of a fishery is the sum of catches from 
local areas. We will explore how details, captured by the copula, of the 
nature of the dependence between local quantities can influence the spatial 
mean or sum and how this may influence human concerns and 
higher organizational levels in 
ecology. Illustrating the idea here, prior work demonstrated that the overall 
reliability of wind-generated electricity depends sensitively on details
of the dependence between wind speeds in multiple wind sites [@li2013]. 
Spatially aggregated ecological variables may show a similar dependence, with
potentially substantial consequences. For instance, if populations of a 
pest species in different locations are all mutually strongly
related to each other in their right tails, then local outbreaks tend strongly
to occur as part of regional epidemics. Similarly, if populations
of a crop resource species are strongly related in their left tails, then 
episodes of local scarcity tend to happen simultaneously, creating regional 
shortages.

<!--DAN: edit and add references to the following-->
Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data. 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between variables; many of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
obtained from A (respectively, B) via monotonic functions that transformed the standard-
normal marginal distributions of A and B to gamma (C, D) or 
uniform (E, F) marginals (details in Appendix X). (G, H) Contours 
of log-scale density functions of Clayton (G) and survival-Clayton 
(H) copulas with parameter 5 (see Methods). (I,J) Reflections
of E, F, respectively, about a vertical axis. Subscripts h and v 
stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
\end{center}
\end{figure}
<!--***DAN: Dan to go back and add an appendix with details as it says. Make sure
to mention the shape and scale params of the gamma marginals in C, D were 2-->




Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure than summarizing in just a single number, the linear correlation coefficient. It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables. If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower tail-dependence" as shown in the first panel of Fig. \ref{fig_cop_pedag} copula plot (right). Similarly, an "upper tail-dependence" copula indicates stronger correlation between higher extreme variables in the raw data. Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution. For example, in the middle panel of Fig. \ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly upper tail dependent (lower tail - upper tail dependence being negative) to dominantly lower tail dependent (lower tail - upper tail dependence being positive) as we move along the row. Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$). So, the flexibility of copulae allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulae from many existing parametric copula families. <!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), described in the \nameref{Methods}. 




From the ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously. Again, if abundance of a species (say, species A) is associated with scarcity of another species (say, species B) then this may indicate causality, i.e. negative feedback of A on species B. If one variable is an environmental factor (like temperatute or rainfall) and another is an observable (like abundance of a species) then copula structure can also help explain the causal relationship between them. On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze the "spatial data" using pairwise bivariate copulae and find if there is any tail dependence or not. If the copulae have asymmetric tail dependence, then it can be viewed as a new way to quantify "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology. In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results. 

We want to address the following questions. \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? \textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure? And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Data}, we described briefly each of bivariate and multivariate datasets we used to answer these questions and we provide a detailed descriptions of the implemented methods in \nameref{Methods}. In the \nameref{Results} section, we summarized our findings for each dataset. The next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related fields with applications shown in real data as well as some simulations. Finally, we conclude in the \nameref{Discussion} section. 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=4 cm,height=4 cm]{./Results/rawdata.pdf}
\includegraphics[width=4 cm,height=4 cm]{./Results/copdata.pdf}\\
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption{First panel : raw data plot has non-uniform (gamma distributed) marginal whereas copula makes its marginal uniform and indicating lower tail dependence in joint distribution, Second and third panel : pedagogical figure showing BB1 copula with varying tail-dependence difference but of same Kendall's tau along each row and with same tail-dependence difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

<!--# Material and methods \label{Material&Methods}-->

# Data \label{Data} 

```{r read_somedata, echo=F, results='hide'}
d_soilCN<-readRDS("Data/dataForShymolinaCoord.rds")
d_soilCN<-na.omit(d_soilCN)
d_birdsBMR<-readRDS("./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")
d_mammalsBMR<-readRDS("./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")
d_cc<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
```

We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data. Copula structure describing the relationship between atmospheric weather variables such as X and Y has been examined previously in the meteorological literature (e.g., REF). We therefore examined environmental variables from the soil instead, using the Rapid Carbon Assessment database (RaCA; @Wills2014).
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Measurments were from the central pedon of a soil core, to 1 m depth.
<!--***DAN: Terry to check the wording and accuracy here-->
Species-trait data were average species basal metabolic rate (KJ per hour) and body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
Survey suction-trap dataset [@Macaulay1988, @Harrington2014]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years; these are 
indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as 
that of @Sheppard2016,
but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 
years data.  Our second population dataset comprised average annual abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological 
Association of the UK. Data pre-processing was similar but not identical to that 
of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
<!--***DAN: Shya, please fill in X with an autolink
***Shya: I mention only number of common plots taken from both biomass and percent cover datasets.-->
plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
the year 2000 [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. 
Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
See Appendix \ref{SM-Data_details} for additional data details. 

The environmental, species-trait, and community datasets, as we studied them, are henceforth 
called the "bivariate" datasets in the sense that they comprise two quantities measured at 
different locations or for different species. The phenology and population datasets are 
"multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
copula structure can be studied for each pair of locations. Likewise the the ecosystem 
functioning (methane flux) dataset was  multivariate in this sense.

# Concepts for Q1: Statistical tools for detecting copula structure \label{Methods}

Question Q1 from the Introduction was answered via statistical analysis of the
data described above, using both model selection and nonparametric approaches.

## Model selection approach \label{Model_selection}

A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) <!--and Bayesian Information Criterion (BIC)--> was used to help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}. For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$ such as the soil C and N data, model selection involved several steps, 
most implemented using tools from the `VineCopula` package. <!--Code combining these tools is in the
[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. to be added in suppmat-->First, we produced ranks 
$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with these "normalized ranked variables" $u_i$ 
and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the smallest element, up to $n$, for the largest. Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
selection algorithms are ineffective if data could not be distinguished from independent data 
since most or all bivariate copula families considered include the independent copula in the 
family or as a boundary case. <!--The independence test was as implemented in the `BiCopIndTest`
function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic. to be added in suppmat-->If independence could be rejected ($0.05$ significance level), model selection proceeded. The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae. These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F). AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2006_copula ; @joe2014_dependence]  and in `VineCopula` documentation. The lower- and upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`. Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

Model selection methods will give the *relative* support of several models, but will not,
on their own, indicate whether any of the models selected is an objectively good fit (all 
models considered might poorly describe the data). To test this, and as the fourth step of
our procedure, we tested the goodness of fit of our best-AIC copula family using a 
bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
two tests, one based on a Cramer-von Mises statistic, and one based on a 
Kolmogorov-Smirnov statistic. For computational efficiency, 
an initial run using $100$ bootstraps was performed, and if
the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
families with that of the Normal family. <!--The same was also done using BIC.....no, we didn't report it-->
The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
the model-averaged lower- and upper-tail dependence
values to the lower- and upper-tail dependence values of a Normal copula, both of which 
are 0. The third part of \hyperref[Intro]{\textbf{(Q1)}}
was answered by comparing the difference of the model-averaged lower- and upper-tail 
dependence values to the analogous difference for a Normal copula, again 0.

For multivariate datasets, e.g., the aphid abundance or first flight data, 
each of which consists of time series from 10 selected locations, we carried out the same bivariate analysis
for all possible pairwise combinations of those locations and compared AIC-weights and tail-dependence of best-fit(minimum AIC) copula with that of the Normal copula. We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa. Then, we reported as the percentage of location pairs showing stronger lower tail dependence out of total non-independent location pairs considered.

## Nonparametric analysis of tail dependence \label{NPA}

In addition to the model selection approach described above, **nonparametric approaches** to quantify lower <!--($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics)--> and upper tail dependence <!--($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics)--> were also used to answer \hyperref[Intro]{\textbf{(Q1)}}. Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust. 
 
Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
are related to each other in the left and right tails, respectively, of their 
distributions [@nelsen2006_copula, @joe2014_dependence]. 
Our nonparametric study of tail dependence uses three
statistics (i.e., $\cor$, $\Ps$ and $\Dsq$) which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions. 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
\end{center}
\end{figure}


The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and sample variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
of $h$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then defined
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
\end{equation}
As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dsq_{l_b,u_b}$ indicate strong dependence. 

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula [@nelsen2006_copula].
The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$. We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulae (for details see Appendix \ref{SM-Test_npa_stat}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no tail dependence (for details see Appendix \ref{SM-surrog_test}). The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
but had Normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were Normal, but the data were otherwise 
statistically unchanged. 

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
from the \nameref{Intro} for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula. 

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets. For multivariate datasets, we used a spatial resampling scheme (for details see Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
approach described above.

# Results for \hyperref[Intro]{\textbf{(Q1)}}: Ecological datasets often have non-Normal copula structures, including asymmetric tail dependence \label{Results}
<!--## question 1
- Do datasets in ecology and related fields have non-Normal copula structure?
- Do they show tail dependence distinct from that of a Normal copula?
- In particular do they show asymmetric tail dependence?

To answer \hyperref[Intro]{\textbf{(Q1)}}, we chose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
--> 

## Results for bivariate data \label{Result_biv}

We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig. \ref{fig_biv_multi_raw_cop}(A-H). For all $4$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Mammals' BMR, Ceder-creek) in Table \ref{tab_bivar_summary}, data were always significantly correlated as indicated by p-value (<0.05) from `BiCopIndTest`{`VineCopula`} (1$^{st}$ row of Table \ref{tab_bivar_summary}). Non-Normal copulae always emerged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}). The Normal copula was found as poor fit compared to highest AIC-weight based asymmetric copulae (3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. Goodness of fit of the best-fit asymmetric copulae were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}, respectively).
Among 4 datasets presented there, first three have significantly stronger upper tail dependence whereas the last one has significantly stronger lower tail dependence compared to Normal (compare 7$^{th}$ and 8$^{th}$ rows of Table \ref{tab_bivar_summary}). It was indicated by negative AIC-weights model averaged (lower-upper) tail dependence (9$^{th}$ row of Table \ref{tab_bivar_summary}) for first 3 datasets and positive AIC-weights model averaged 
(lower-upper) tail dependence for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}} (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}). We also reported the rank of ($\Ps_l - \Ps_u$) for the actual copula data in camparison to the same non-parametric statistics of Kendall-correlation preserving 1000 Normal surrogates. Ranks if were $>$ 975 or $<$ 25, then indicated a significantly (95$\%$ CI) different lower tail-dependence or upper tail-dependence measures, respectively, based on Normal-null hypothesis. We found our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}). 

### Example for bivariate case : soil C and N data shows upper tail dependence\label{Ex_SoilCN}

```{r read_BivMS_soilCN_data,echo=F}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

We will now discuss briefly findings from one of our bivariate datasets as follows. As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

<!--#### Model selection approach \label{Ex_SoilCN_Model_selection}-->

From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`). We then fitted several copulae (all $16$ families as mentioned in \nameref{Model_selection})
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulae (Table \ref{SM-tab_soilCNfit}). 
Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit. The table shows that several copulae are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data. 

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data. Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. Therefore, we also verified this finding with \nameref{NPA} approach (see Fig. \ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics. Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.

Furthermore, the difference of model-averaged tail dependence statistics (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}. To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).

<!--#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}-->

## Results for multivariate spatial data \label{Result_multi}
```{r read_RES_aphid_count, echo=F}
RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
```

We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other. Here, to present our results we chose one species from each category. We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations). 

We found major occurence of non-independent copulae (row 2, Table \ref{tab_multivar_summary}) out of all pairwise locations (row 1, Table \ref{tab_multivar_summary}) considered. Non-Normal copulae, most of the time, emerged as best-fit based on minimum AIC out of 16 copula family considered (row 3, Table \ref{tab_multivar_summary}) answering the first part of question \hyperref[Intro]{\textbf{(Q1)}}. All four data sets showed Normal copulae were always poor fit in comparison to best-fit non-Normal copulae (compare their average AIC-weights from row 4 and row 5 of Table \ref{tab_multivar_summary}). We considered the "good-fit" location pair based on the p-values (>0.01) obtained from both Cramer-von Mises and Kolmogorov-Smirnov statistics (row 6, Table \ref{tab_multivar_summary}). Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}. The spatial synchrony (from model selection) results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile). Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.

### Example for multivariate case : abundance data for the green spruce aphid (*Elatobium abietinum*) shows spatial synchrony with lower tail dependence\label{Ex_count}

<!--#### Model selection approach \label{Ex_count_Model_selection}-->

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulae was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulae was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulae of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100`$\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question \hyperref[Intro]{\textbf{(Q1)}} for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis. 

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`) which indicates aphids are synchronously rare at most of the location pairs. The dominance of lower tail dependence in green spruce aphid's abundance data with asymmetric copula structures answers the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Fig. \ref{fig_biv_multi_raw_cop}(I) and (J) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

<!--#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}-->

As mentioned in \nameref{NPA} and Appendix \ref{SM-spatial_resamp}, mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}. Visual illustration of those results is discussed in Appendix \ref{SM-visual_npa}.

<!--
```{r tab_bivar_summary, echo=F, results='markup'}
tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
knitr::kable(tab_bivar_summary, 
             format='pandoc',
             caption = "Summary of results for bivariate data-sets : \\textbf{p.Independence.test}: p-value from `BiCopIndTest`, \\textbf{minimumAIC.bestfit.copula}: best fit copula based on minimum AIC, \\textbf{Bestfit.copula.AICw}: AIC-weight of that best fit copula, \\textbf{Normal.copula.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence. Statistic $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and ranked in ascending order. Rank for real data \\textbf{Rank.(Pl-Pu)}, if were $>975$ or $<25$, then indicated a significantly ($95\\%$ CI) different lower or upper tail-dependent measures based on Normal-null hypothesis.
\\label{tab_bivar_summary}",
             booktabs=TRUE)
```

```{r tab_multivar_summary, echo=F, results='markup'}

tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
knitr::kable(tab_multivar_summary, 
             format='pandoc',
             caption = "Summary of results for multivariate data-sets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01 
(both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the non-parametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
             booktabs=TRUE)
```
-->

<!--bivariate table summary in latex format-->
```{r tab_bivar_summaryl, echo=F, results='asis',message=F}
tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
row.names(tab_bivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(Description = c("1. p-value from independence test",
           "2. Best fit copula based on minimum AIC",
           "3. AIC-weight of that best fit copula",
           "4. AIC-weight of Normal copula",
           "5. p-value from goodness of fit test for Cramer-von Mises method",
           "6. p-value from goodness of fit test for Kolmogorov-Smirnov method",
           "7. Model averaged (AIC-weights based) lower tail dependence",
           "8. Model averaged (AIC-weights based) upper tail dependence",
           "9. Model averaged (AIC-weights based) (lower - upper) tail dependence",
           "10. Rank of statistics (Pl-Pu) for real data in comparison to 1000 Normal surrogates")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
knitr::kable(tab_bivar_summary, 
             format="latex", 
             caption = "Summary of results for bivariate data-sets\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "SoilCN", "Birds BMR", "Mammals BMR", "Ceder-creek")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "11cm")%>%kable_styling(latex_options = "basic",position = "center")
```

<!--multivariate table summary in latex format-->

```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
tab1<-tab_multivar_summary[c(1:7,9,11,12,14,16),] # see carefully if you rearrange tab_multivar_summary's row
                                                   #in suppmat this will change
tab2<-tab_multivar_summary[c(8,10,13,15,17),]  # see carefully if you rearrange tab_multivar_summary's row
                                                   #in suppmat this will change

two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[c(7:8,10:12),i], " , ",tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change

colnames(two_tab_in1)<-colnames(tab_multivar_summary)
rownames(two_tab_in1)<-c("CI_LT","CI_UT","CI_Pl","CI_Pu","CI_Pl-Pu")

tab_multivar_summary<-rbind(tab_multivar_summary[c(1:6),],
                            two_tab_in1[1:2,], #CI of LT, CI of UT
                            tab_multivar_summary[11,],
                            two_tab_in1[5,] #CI of Pl-Pu
                           )

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Total number of location pairs (excluding self-comparison)",
           "2. Number of location pairs having non-independent copula structures",
           "3. Number of non-Normal copula as best fit",
           "4. Average AIC-weights of best fit copula from all location pair",
           "5. Average AIC-weights of Normal copula from all location pair",
           "6. Number of good-fit location pair*",
           "7. 95% CI of model-avg. lower tail-dep. from all location pair",
           "8. 95% CI of model-avg. upper tail-dep. from all location pair",
           "9. Percentage of location pairs showing stronger lower tail dependence",
           "10. 95% CI of (Pl - Pu) from all location pair")
           )

tab_multivar_summary<-cbind(dt,tab_multivar_summary)

knitr::kable(tab_multivar_summary, 
             format="latex", 
             caption = "Summary of results for multivariate data-sets\\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green-spruce aphid's count",
                        "Leaf-plum curling aphid's firstflight",
                        "Ceratium furca's count","Methane-flux")
             )%>%column_spec(2:5,width="2.5cm")%>%
             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")%>%
   add_footnote("Good fit was considered for best fit copula with p-values > 0.01 (both Cramer-von Mises and Kolmogorov-Smirnov method)", notation="symbol", threeparttable = T, escape = F)
```

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{7 cm} (I) \hspace{3.2 cm} (J)} \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{7 cm} (K) \hspace{3.2 cm} (L)} \\
\vspace{-0.6 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs. log(mass) for birds, \textbf{(C)} log(BMR) vs. log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. Lower panel shows \textbf{(I - J)} raw data plots and \textbf{(K - L)} copula plots for  green spruce aphid's count data from location pair(4,11) and for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}


<!--Question 2 analyses-->
# Concepts and models for \hyperref[Intro]{\textbf{(Q2)}}: possible causes of non-Normal copula structure \label{Causes}

This section is dedicated to exploring question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. We showed environmental copula structure and asymmetric sensitivities to the environment could produce population copula structures.
<!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?--> 
<!--Moran effects-->
\textbf{(i)} We were interested to extend the basic idea of Moran's theorem : if environmental noises at two different habitat patches had some specific assymetric tail dependence then what will be the outcome dependence in the populations at those habitat patches. Should their dependence be exactly the same as of the noises or not ? To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq. \ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.  

\begin{equation}\label{eq.AR1}
P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
\end{equation}

To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$. We showed with Eq. \ref{eq.AR1} in the next section that both model selection method and non-parametric analysis indicated similar tail dependence in population copula as that of noise copula. <!--Asymmetric sensitivity and copula structure in populations-->
\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq. \ref{eq.AS_g}). To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}, Fig. \ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$. Local environmental fluctuation function $h(\delta)$ varied according to Eq. \ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$. $\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$. With this model we showed tail dependence in population copula changed from lower to upper tail dependence as global noise copula $g(\varepsilon)$ changed its form from $g_{1}$ to $g_{2}$ (Eq. \ref{eq.AS_g}).

\begin{align}
g(\varepsilon) = \begin{cases} 
g_{1} = \begin{cases}
\varepsilon & \text{if $\varepsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\varepsilon & \text{if $\varepsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

\begin{equation}\label{eq.AS_h}
h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
\end{equation}

# Results for \hyperref[Intro]{\textbf{(Q2)}}: similar tail dependence in population copula as that of noise copula \label{Results_Q2}

To test environmental effect on population copula from the model \ref{eq.AR1}, our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2006_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively. Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig. \ref{fig_Cause4copula_params_GOF_500}). We also tested the goodness of fit of the estimated parameters for the population-copula and its standard error due to sample size. We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov). p-values greater or equal to $0.05$ (above the dashed line in Fig. \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters. Fig. \ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig. \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$. (For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function). 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit (not significantly different) of population copula with same noise copula family. \label{fig_Cause4copula_params_GOF_500}}
\end{center}
\end{figure}

Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations. Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family. Fig. \ref{SM-fig_Cause4copula_C_kend} compares the statistics from population copula and the copula made of noises drawn from lower tail dependent Clayton family. It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g. see Fig. \ref{SM-fig_Cause4copula_SC_spear}). We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which verified the Moran effect on the population data caused by environmental noise. (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function). Linear regression between tail statistics from noise and population copula (Fig. \ref{fig_Cause4copula_scatter_C_kend}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig. \ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula. Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend}}
\end{center}
\end{figure}


To explore the effect of asymmetric sensitivity, we simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$. $10000$ points were generated and we kept only the last $1000$ points to plot the population-copula (Fig. \ref{fig_ms_asym_b_0.1}) for each specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function). When asymmetric global noise had only non-positive contributions at two habitat patches ($g_{1}$ in the first panel of Fig. \ref{fig_ms_asym_b_0.1}), lower tail dependence appeared in the global noise copula and so thus in affected population copula (last subplot in the first panel of Fig. \ref{fig_ms_asym_b_0.1}). Similarly, when global noises had only non-negative contributions at two habitat patches ($g_{2}$ in the second panel of Fig. \ref{fig_ms_asym_b_0.1}), upper tail dependence appeared in the global noise copula and so thus in the outcome population copula (last subplot in the second panel of Fig. \ref{fig_ms_asym_b_0.1}). We also carried out model selection and non-parametric statistical analysis with each copula of Fig. \ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_2}. According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations. Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach. 

\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{Effects of asymmetric global noise copula (resulting from the asymmetric dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq. \ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we dont really know how a change in the environment copula structure would make what kind of change in the population copula structure.--> 
<!--
## Application : The Moran effect and copula structure in population 

```{r read_RES_temp_aphid, echo=F}
RES_temp_aphid<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_winter_neg_temp.RDS")
```

Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables. This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations. In the previous section, we already tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises. Here, we investigated the probable cause of upper tail dependence in the leaf-curling plum aphid's firstflight date throughout all the locations. We assumed that average winter temperature which might be spatially synchronized and had asymmetric tail dependence cause the synchronized delay in leaf-plum curling aphid's appearence. We tested our hypothesis and indeed found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 35 years (1976-2010). We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data. Negative winter temperature was considered for simulation to compare the aphid's firstflight results with temparature copula (aphid's firstflight date were negatively correlated with temperature or in other words positively correlated with negative temparature). Model selection results for negative average winter temperature in 10 $\times$ 10 matrices were represented in Table \ref{SM-tab_temp_aphid_bestcop} - Table \ref{SM-tab_temp_aphid_LTmUT}. In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data. Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulae with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}). -->

### Need to think about mechanisms for biogeochemical, evolutionary, and community data

<!--Question 3 analyses-->
# Concepts for \hyperref[Intro]{\textbf{(Q3)}}: Consequences of non-Normal copula structure and tail dependence for our understanding of biology \label{Consequences}

This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We wanted to examine whether the asymmetric tail dependence had any **(i)** effects on skewness of spatially averaged time series data. We predicted that because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, should change significantly from the individual skewness of time-series data. We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set in the next section (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)). We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not. If the $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).

We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}(I)). In the following section we will show application of these ideas with real datasets.
<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-----------
Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}). We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion). For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig. \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig. \ref{SM-fig_risk_K_effect}). For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one. For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster. -->

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid firstflight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation). \label{fig_skewness_spearman}}
\end{center}
\end{figure}

## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of skewness

We used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data. We carried out the analysis mentioned in the concepts above, with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species. In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal. For example, Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates. Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the green-spruce aphid count data came from a normal copula. In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data. As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}). On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was significantly higher than expected if the leaf-plum curling aphid first flight data came from a normal copula, which was also confirmed by upper-tail dependence results from model selection and non-parametric results (see second column of Table \ref{tab_multivar_summary}). Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula. Our previous results on non-parametric analysis and model fitting also supported lower tail dependence (see third column of Table \ref{tab_multivar_summary}). But, as $p>0.05$, we could not reject $H_{0}$ in that case. Finally, for methane-flux data, the averaged skewness was found significantly lower in comparison to its Normal surrogates (Figure \ref{fig_skewness_spearman} D). We found the same result with all datasets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_kendall}).


## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of curvature of spatial Taylor's law regression line

```{r read_tl_aphid_count_sp6, echo=F}
tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
```

Spatial Taylor's law, the observed linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017]. We already showed through simulation (Appendix \ref{SM-Consequence_TL_simulation}) that for asymmetric copula this linear relationship somehow deviates compared to Normal copula for a distribution with gamma marginals. We tested spatial Taylor's law with aphid count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs. `log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $25\%$ of total species considered. For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and non-parametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}). 

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $9961$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data). $(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for corn-leaf aphid abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean). \label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}

# Discussion \label{Discussion}

<!--what we did and why is it important to study?-->

Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables. We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selction results. Our approach could help to interpret the synchronous behavior of a species from several metapopulations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongerly correlated than lower values. 

We explained the possible causes which induced asymmetric tail-dependence in the copula. Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig. \ref{fig_Cause4copula_params_GOF_500} and Fig. \ref{fig_Cause4copula_scatter_C_kend}). We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as the initial global noise copula (Fig. \ref{fig_ms_asym_b_0.1}). We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula. For example, Fig. \ref{fig_skewness_spearman}(A) showed green spruce aphid abundance timeseries had right skewed distribution but its skewness was signifiantly lower than was expected if it came from Spearman correlation preserving Normal surrogates. This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure. Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emmisions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment. 

<!--future work-->
We think this copula approach have scope to investigate into the extinction risks of metapopulation due to climate change. To illustrate, lower-tail dependent environmental fluctuations may suppress the metapopulations synchronously by making some of the habitats go extinct. This could help in managing policy for conservation of endangered populations.  


\clearpage

# References


