---
title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, other authors, Daniel Reuman"
date: "February 18, 2018"
geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty
      
tables: True
link-citations: True
urlcolor : blue

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---
updated on `r Sys.Date()`
```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
#seed<-101
#source("mtime.R") #A function needed for caching
```
# Front matter

*Affiliations:*

Reuman: 

Ghosh:

*Possible alternative titles:*

- A complete statistical description of dependence between variables for ecology and related fields

*Keywords:*

*Type of Article*

- Ideas and Perspectives

*The number of words in the abstract*

*The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)* 

*The number of words in each text box*

*The number of references*

*The number of figures, tables, and text boxes*

*The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent* 

\newpage
# Abstract

\newpage
# Introduction

*Goes in the last paragraph of the intro:*
We address the following questions. 1) Do datasets in ecology and related fields have non-Gausian copula srtructure? Do they show tail dependence distinct from that of a Gaussian copula, and in particular do they show asymmetric tail dependence? 2) If so, what are some possible causes of non-Gaussian copula structure? And 3) what are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?

# Material and methods

## Datasets


## Statistical methods  \label{Methods}

### Model selection approach \label{Model_selection}

For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
most implemented using tools from the `VineCopula` package. Code combining these tools is in the
[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 

First, we produced ranks 
$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
smallest element, up to $n$, for the largest. 

Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
selection algorithms are ineffective if data could not be distinguished from independent data 
since most or all bivariate copula families considered include the independent copula in the 
family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic. If independence could be rejected ($0.05$ significance level), model selection proceeded.

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

Model selection methods will give the *relative* support of several models, but will not,
on their own, indicate whether any of the models selected is an objectively good fit (all 
models considered might poorly describe the data). To test this, and as the fourth step of
our procedure, we tested the goodness of fit of our best-AIC copula family using a 
bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
two tests, one based on a Cramer-von Mises statistic, and one based on a 
Kolmogorov-Smirnov statistic. For computational efficiency, 
an initial run using $100$ bootstraps was performed, and if
the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

<!--Shya please insert some additional material for multivariate datasets - perhaps
you don't need much because basically you just ran the bivariate analysis on each pair
of locations, but you need to say that and then maybe a few sentences about how 
everything was tallied up after having done that ??? -->

### Non-parametric analysis of tail dependence \label{NPA}

Left- and right-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
are related in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}(A)). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions. 

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{5.6 cm} (B)} \\
\vspace{-0.6 cm}
\begin{center}
\hspace{0.5 cm}
\includegraphics[width=5 cm,height=4.3 cm]{./stat_image.jpg} \hspace{0.6 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/CorlmCoru_C.pdf}
\end{center}
\textbf{ \hspace{6 cm} (C) \hspace{5.7 cm} (D)} \\
\vspace{-0.6 cm}
\begin{center}
\hspace{0.2 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/PlmPu_C.pdf} \hspace{0.3 cm}
\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/D2umD2l_C.pdf}
\caption{(A) Schematic diagram supporting defintions of nonparametric tail dependence statistics; (B) Results of tests of $\cor_l-\cor_u$, (C) $\Ps_l-\Ps_u$, and (D) $\Dsq_u-\Dsq_l$ reveal the known asymmetry (more lower tail than upper) of tail dependence of the Clayton copula with $n=35$ data points. \label{fig_stat_image}}
\end{center}
\end{figure}

The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
of $h$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then defined
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
\end{equation}
As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dsq_{l_b,u_b}$ indicate strong dependence. 

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula [@nelsen2007_copula].
The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas. Fig. \ref{fig_stat_image}(B - D) represent the results for Clayton copulas with $n=35$ data points and $100$ simulations (for details see Appendix \ref{SM-Test_npa_stat}). 

For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same statistics computed on surrogate datasets that were produced from the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the empirical data no different from what would have been expected if the copula structure of the data were normal, but the data were otherwise statistically unchanged.
For bivariate datasets, surrogates were produced as follows.   

- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
study using the `iTau` function of the `VineCopula` package). 
- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
$\mathcal{B}(\theta_\tau)$. 
- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
indices $1,\ldots,n$). 
- Likewise permute $(y_1,\ldots,y_n)$
such that the permutation
$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
another permutation of $1,\ldots,n$).
- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$. 

Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
[`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 

Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.  

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
provides a test of asymmetry of tail dependence. 

For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
The scheme is identical to that proposed by [@bjornstadfalck2001] for application in their now 
widely used spline correlogram method; the same scheme was also used by [@walteretal2017] as part of a method for 
performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017]. 

<!--Dan to insert a description of the surrogate approach here.-->


# Results

## question 1

- Do datasets in ecology and related fields have non-Gausian copula structure?
- Do they show tail dependence distinct from that of a Gaussian copula?
- In particular do they show asymmetric tail dependence?

### soil C and N data \label{Ex_SoilCN}

```{r read_BivMS_soilCN_data,echo=F}
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

#### Model selection approach \label{Ex_SoilCN_Model_selection}

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1. 

Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}

As described in \nameref{Methods}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). 
<!--
Because the correlation 
used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
should be used, so the top-right panel of Fig. \ref{SM-fig_soilCN_nonparam}
is more reliable than the top left.
--> 
Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

### BMR data \label{Ex_BMR}

```{r read_BivMS_BMR_data,echo=F}
BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
```

#### Model selection approach \label{Ex_BMR_Model_selection}

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}). The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence. We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}

As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig. \ref{SM-fig_BMR_nonparam}). 

Results indicate that dependence of data is different in some bins than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.

### Anolis data \label{Ex_contrasts}

```{r read_BivMS_contrasts_67_data,echo=F}
BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

#### Model selection approach \label{Ex_contrasts_Model_selection}

The test of indepedence gave $p=0$ <!--`r BivMS_contrasts_67$IndepTestRes` this is e-15 could not compile latex--> (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}). The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1. 

These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}

As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins). Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}). 

Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. 

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B). Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

```{r read_RES_aphid_count, echo=F}
RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
```

#### Model selection approach \label{Ex_count_Model_selection}

As described in Methods, model selection among copulas was done separately for each pair of the 10 
sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
to the AIC weight 
of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data. Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis. 

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1. 

#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
structure do not have, providing further evidence answering the third part of question 1. 

Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with 
distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig. \ref{SM-fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig. \ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
were significantly greater than zero, demonstrating asymmetry of tail dependence. Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package. 

### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}

```{r read_RES_aphid_ff,echo=F}
RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
```

#### Model selection approach \label{Ex_ff_Model_selection}

As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
and columns corresponding to locations. Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families. Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the 
hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}). In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
the model-averaged (again, AIC based) upper-tail dependence statistic are in 
Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis. 

The difference between the model-averaged lower-tail dependence statistic and the 
model-averaged upper-tail dependence statistic (lower minus upper), shown in 
Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1. We note this result is opposite the analogous result for aphid abundance
data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence. 

#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
zero and mean values were positive, indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating. 

Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance. Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig. \ref{SM-fig_aphid_ff} C-F).
$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence. 

### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

```{r read_RES_plankton_north_sea,echo=F}
RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
```

#### Model selection approach \label{Ex_pns_Model_selection}

As described in Methods and as in previous sections, model selection among copulas was done separately for 
each pair of sampling locations. There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations. Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal}) 
confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}). In all cases at least
`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
location pairs and negative for others, but was positive for 
`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1. 

#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
always excluded $0$ and mean values were positive, indicating tail dependence.
Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
structure do not have, providing further evidence answering the third part of question 1.

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig. \ref{SM-fig_pns} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig. \ref{SM-fig_pns} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the 
statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
quantities across all pairs of sampling locations were significantly greater 
than zero, showing asymmetry of tail dependence.

### Methane flux data \label{Ex_methane}

```{r read_RES_methane,echo=F}
RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
```

#### Model selection approach \label{Ex_methane_Model_selection}

As described in Methods and as in previous sections, model selection among copulas was done separately for 
each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}). In all cases at least
`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
location pairs and negative for others, but was positive for 
`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1. 

#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
always excluded $0$ and mean values were positive, indicating tail dependence. But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig. \ref{SM-fig_methane} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
their opposite meaning (Fig. \ref{SM-fig_methane} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the 
statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.

<!--Question 2 analyses-->
## Some possible causes of non-Gaussian copula structure 

Q1/2: How are copula structure of ecologically influential climate variables changing as part of climate change? We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the pops, but we don’t really know how a change in the environment cop structure would make what kind of change in the pop cop structure. How?

### The Moran effect and copula structure in populations 

We show here that if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables.

To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq. \ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.  

\begin{equation}\label{eq.AR1}
P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
\end{equation}

To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$. We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches. Should their dependence be exactly the same as of the noises or not ? 

To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively. Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig. \ref{SM-fig_Cause4copula_params_GOF_500}). We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov). p-values greater or equal to $0.05$ (above the dashed line in Fig. \ref{SM-fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters. Fig. \ref{SM-fig_Cause4copula_params_GOF_2500} represents similar figure as of Fig. \ref{SM-fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$. (For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function). 

Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations. Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family. Fig. \ref{SM-fig_Cause4copula_C_spear} and Fig. \ref{SM-fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig. \ref{SM-fig_Cause4copula_SC_spear} and Fig. \ref{SM-fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family. We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise. (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function). 

### Asymmetric sensitivity and copula structure in populations
<!--Another model, distinct from the above, Dan and Shya to discuss-->
Here we intend to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) beyond a certain expectation threshold. To test this assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;

\begin{equation}\label{eq.AS}
P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
\end{equation}

$b$ was the linear growth rate. Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq. \ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.

\begin{align}
g(\epsilon) = \begin{cases} 
g_{1} = \begin{cases}
\epsilon & \text{if $\epsilon< 0$}; \\
0 & \text{otherwise}.
\end{cases}\\ \\
g_{2} = \begin{cases}
\epsilon & \text{if $\epsilon> 0$}; \\
0 & \text{otherwise}\label{eq.AS_g}.
\end{cases}
\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
\end{align}

Local environmental fluctuation function $h(\delta)$ varied according to Eq. \ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$. $\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.

\begin{equation}\label{eq.AS_h}
h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
\end{equation}

We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$. $10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulation to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function). Fig. \ref{SM-fig_ms_asym_b_0.1} and Fig. \ref{SM-fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively. We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_8}. Table \ref{SM-tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test. In each case, successful bootstrap was $100\%$. According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best. It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently. Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_8}) are consistent with this model selection approach. 


### Need to think about mechanisms for biogeochemical, evolutionary, and community data

<!--Question 3 analyses-->
## Consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?

### Effect on skewness of the spatial mean

Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location. Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness. We tested this with null hypothesis $H_{0}$. If $H_{0}$ is true that means our multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula. So, first we generated $10000$ number of surrogate normal copula from our given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same. Then we plotted the distribution of spatial mean of each surrogate time series data set and the given multivariate data. If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence.

To verify that we used aphid-count data, aphid-firstflight data, plankton data and methane-flux data. For aphid data we choose only those locations for which atleast $30$ number of data points present in each time series when taken as pairwise locations. For plankton-UK-sea data and methane-flux data that threshold on number of data points are $45$ and $50$, respectively. 

- Aphid count \label{skewness_count}

Figure \ref{SM-fig_skewness_count} shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula. Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_count}). As $p<0.05$, we reject $H_{0}$. This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence. 

- Aphid first flight \label{skewness_ff}

Figure \ref{SM-fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula. Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}). As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence. 


- UK Sea plankton \label{skewness_pns}

Figure \ref{SM-fig_skewness_pns} shows that averaged skewness is lower than expected if the plankton data came from a normal copula. Our previous results on non-parametric analysis and model fitting showed lower tail dependence. Here, as $p>0.05$, we cannot reject $H_{0}$, .......???

- Methane-flux data \label{skewness_methane}

Figure \ref{SM-fig_skewness_methane} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula. Our previous results on non-parametric analysis and model fitting showed lower tail dependence. As $p<0.05$, we reject $H_{0}$. This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence. 

### Effect on extinction risk

In order to study the effects of non-gaussian copula structure on extinction risk for population of a species at different locations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
<!--did I write correct spelling for lewton model?-->

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which may take strong right tail (or left tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix. $D$ can take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch can interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;


- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first plot of Fig. \ref{SM-fig_risk_local_disp} and Fig. \ref{SM-fig_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations. [see Fig. \ref{SM-fig_risk_local_disp} - \ref{SM-fig_risk_global_disp}]
 
    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{SM-fig_risk_K_effect}]
  
With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig. \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig. \ref{SM-fig_risk_numlocs_effect_global_disp}] for both model. 



# Discussion



\newpage
# References


