---
title: "Supporting Information"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 12, 2018"
output: pdf_document
---

# Descriptions of data


# Testing the ranking and bivariate copula display function

This is just for testing the `getcopula` function.

<!--Source in function and a package--->
```{r}
source("getcopula.R")
require(copula)
require(mvtnorm)
```

Plot data from a Clayton copula, which has lower tail dependence:
```{r}
copC<-claytonCopula(8,2)
dC<-rCopula(1000,copC)
vC<-getcopula(d=dC,rankon=F,ploton=T)
```

Plot data from a Joe copula, which has upper tail dependence:
```{r}
copJ<-joeCopula(8,2)
dJ<-rCopula(1000,copJ)
vJ<-getcopula(d=dJ,rankon=F,ploton=T)
```

Get data from a bivariate normal distribution and take ranks and
then plot:
```{r}
sigma<-matrix(c(1,.7,.7,1),2,2)
dN<-rmvnorm(1000,mean=c(0,0),sigma=sigma)
vN<-getcopula(d=dN,rankon=T,ploton=T)
```

# Testing the bivariate model selection and goodness-of-fit pipeline

This is just for testing the `OurBiCopSelect` function. We test that
it runs and gives reasonable results by simulating data and seeing if 
you recover what kind of copula was used.

<!--Source in the function and a package, changeable constants-->
```{r}
source("OurBiCopSelect.R")
require(copula)
nbsmall<-100
pthresh<-0.2
nblarge<-1000 #for bootstrapping, see below
numdat<-500 #amount of artificial data
seed<-101
```

<!--A function needed for caching-->
```{r echo=F}
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}
```

If real data come from a Clayton copula, independence should
be rejected, Clayton should be the AIC best, it should pass the
goodness of fit test, and the normal copula should fail the
goodness of fit test:
<!--Test for real data from a clayton-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

families<-c(1,3:6,13,14,16)

#data are really from a clayton
cop<-claytonCopula(5,2)
dat<-rCopula(numdat,cop)
allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresC
```

Same test, but data now come from a Gumbel copula:
<!--Test for real data from a gumbel-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a gumbel
cop<-gumbelCopula(6,2)
dat<-rCopula(numdat,cop)
allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresG
```

Same test, but data now come from a Frank copula:
<!--Test for real data from a frank-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a frank
cop<-frankCopula(5,2)
dat<-rCopula(numdat,cop)
allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresF
```

Same test, but data now come from a Joe copula:
<!--Test for real data from a joe-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a joe
cop<-joeCopula(6,2)
dat<-rCopula(numdat,cop)
allresJ<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresJ
```

# Testing the pipeline for examining tail dependence in bivariate data

